---
title: "Amandafw_OriginalHomeworkCode_03"
author: "Amanda Wu"
date: "2023-10-09"
output: html_document
---

##uploading data into R
```{r Uploading Data}
library(curl)
f <- curl("https://github.com/fuzzyatelin/fuzzyatelin.github.io/blob/master/AN588_Fall23/zombies.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE) #this is the typical path used in challenges but would give an error for me - "Error in open.connection(file, "rt")" or another error with the final code line
#or 
d <- read.csv("zombies.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE) #this seemed to be more effective
head(d)
summary(d)
```

## Question 1
```{r Question 1}
#Calculate the population mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, and years of education). NOTE: You will not want to use the built in var() and sd() commands as these are for samples.
names(d)
mean(d$height, na.rm = TRUE) #the mean for heights
h <- d$height
sqrt(sum((h - mean(h))^2)/length(h)) #pop standard deviation is the square root of the pop variance; population variance is the sum of squares divided by # of values

mean(d$weight, na.rm = TRUE)
w <- d$weight
sqrt(sum((w - mean(w))^2)/length(w)) #pop sd

mean(d$age, na.rm = TRUE)
a <- d$age
sqrt(sum((a - mean(a))^2)/length(a)) #pop sd

mean(d$zombies_killed, na.rm = TRUE)
z <- d$zombies_killed
sqrt(sum((z - mean(z))^2)/length(z)) #pop sd

mean(d$years_of_education, na.rm = TRUE)
y <- d$years_of_education
sqrt(sum((y - mean(y))^2)/length(y)) #pop sd
```


## Question 2
```{r Question 2}
#Use {ggplot} to make boxplots of each of these variables by gender.
library(ggplot2)
names(d)

h <- ggplot(data = d, aes(x = gender, y = height))
h <- h + geom_boxplot()
h <- h + ylab("Height") + xlab("Gender")

w <- ggplot(data = d, aes(x = gender, y = weight))
w <- w + geom_boxplot()
w <- w + ylab("Weight") + xlab("Gender")

a <- ggplot(data = d, aes(x = gender, y = age))
a <- a + geom_boxplot()
a <- a + ylab("Age") + xlab("Gender")

z <- ggplot(data = d, aes(x = gender, y = zombies_killed))
z <- z + geom_boxplot()
z <- z + ylab("Zombies Killed") + xlab("Gender")

y <- ggplot(data = d, aes(x = gender, y = years_of_education))
y <- y + geom_boxplot()
y <- y + ylab("Years of Education") + xlab("Gender")

library("gridExtra") #package i found to present all the graphs together
grid.arrange(h, w, a, z, y, ncol = 2, nrow = 3) 
```
Source: https://www.geeksforgeeks.org/draw-multiple-ggplot2-plots-side-by-side/
Notes: the graphs look a little condensed though and I'm not too sure how to fix it or whether I should change it so each graph has their own space, however, the knitted code looks really long when I do this

## Question 3
```{r Question 3}
#Use {ggplot} to make scatterplots of height and weight in relation to age. Do these variables seem to be related? In what way?
p <- ggplot(data = d, aes(x = height, y = age)) #entering the data
p <- p + xlab("Height") + ylab("Age")  #renaming the axis
p <- p + geom_point()  #make a scatterplot
p <- p + geom_smooth(method = "lm", fullrange = TRUE) #adding regression lines

w <- ggplot(data = d, aes(x = weight, y = age)) #entering the data
w <- w + xlab("Weight") + ylab("Age")  #renaming the axis
w <- w + geom_point()  #make a scatterplot
w <- w + geom_smooth(method = "lm", fullrange = TRUE)

grid.arrange(p, w, ncol = 2)
```
There appears to be a more definitive correlation between height and age than weight and age but both variables generally increase with age. Both weight and height appears to be correlated with taller individuals typically having larger weights. 


## Question 4
```{r Question 4}
#Using histograms and Q-Q plots, check whether the quantitative variables seem to be drawn from a normal distribution. Which seem to be and which do not (hint: not all are drawn from the normal distribution)? For those that are not normal, can you determine from which common distribution they are drawn?
summary(d)
#creating histogram & Q-Q plots with height:
hist(d$height, breaks = seq(from = 50, to = 85, by = 1), probability = TRUE, xlab = "Height", main = "Histogram for Height") #the breaks were determined with summary(d) from the min and the max values
qqnorm(d$height, main = "Q-Q plot for Height")
qqline(d$height, col = "red")

#histogram for weight
hist(d$weight, breaks = seq(from = 85, to = 215, by = 1), probability = TRUE, xlab = "Weight", main = "Histogram for Weight")
qqnorm(d$weight, main = "Q-Q plot for Weight")
qqline(d$weight, col = "red")

#histogram for zombies killed
hist(d$zombies_killed, breaks = seq(from = -5, to = 15, by = 1), probability = TRUE, xlab = "Zombies Killed", main = "Histogram for Zombies Killed")
qqnorm(d$zombies_killed, main = "Q-Q plot for Zombies Killed")
qqline(d$zombies_killed, col = "red")

#histogram for age
hist(d$age, breaks = seq(from = 5, to = 35, by = 1), probability = TRUE, xlab = "Age", main = "Histogram for Age")
qqnorm(d$age, main = "Q-Q plot for Age")
qqline(d$age, col = "red")

#histogram for years of education
hist(d$years_of_education, breaks = seq(from = -5, to = 13, by = 1), probability = TRUE, xlab = "Years of Education", main = "Histogram for Years of Education")
qqnorm(d$years_of_education, main = "Q-Q plot for Years of Education")
qqline(d$years_of_education, col = "red")
```
Height, Weight, and Age appears to be drawn from a normal distribution as they seem to adhere towards the line in the qqplots. In addition, their histograms appear similar to the normal distribution. Years of Education and Zombies Killed did not seem to be drawn from a normal distribution as their qq plots were really different and their histograms appear more similar to poission distributions.


## Question 5
```{r Question 5}
#Now use the sample() function to sample ONE subset of 30 zombie survivors (without replacement) from this population and calculate the mean and sample standard deviation for each variable. Also estimate the standard error for each variable, and construct the 95% confidence interval for each mean. Note that for the variables that are not drawn from the normal distribution, you may need to base your estimate of the CIs on slightly different code than for the normal…
#standard error is the standard deviation of the sampling distribution 
set.seed(1) #makes sure that the sample doesn't change each time
sh <- sample(d$height, 30, replace = FALSE)
m_sh <- mean(sh) #mean
sigma <- sqrt(sum((sh - mean(sh))^2)/length(sh)) #sample standard deviation
sigma
sh_se <- sqrt(sigma^2/30) #standard error
sh_se
upper <- mean(sh) + qnorm(0.975, mean = 0, sd = 1) * sh_se
lower <- mean(sh) + qnorm(0.025, mean = 0, sd = 1) * sh_se  #ci with normal distribution
ci <- c(lower, upper)
ci

sw <- sample(d$weight, 30, replace = FALSE)
m_sw <- mean(sw)
sigma <- sqrt(sum((sw - mean(sw))^2)/length(sw)) 
sw_se <- sqrt(sigma^2/30) 
sw_se
upper <- mean(sw) + qnorm(0.975, mean = 0, sd = 1) * sw_se
lower <- mean(sw) + qnorm(0.025, mean = 0, sd = 1) * sw_se  #ci with normal distribution
ci <- c(lower, upper)
ci

sa <- sample(d$age, 30, replace = FALSE)
mean(sa)
sigma <- sqrt(sum((sa - mean(sa))^2)/length(sa))
sa_se <- sqrt(sigma^2/30) 
sa_se
upper <- mean(sa) + qnorm(0.975, mean = 0, sd = 1) * sa_se
lower <- mean(sa) + qnorm(0.025, mean = 0, sd = 1) * sa_se  #ci with normal distribution
ci <- c(lower, upper)
ci

sz <- sample(d$zombies_killed, 30, replace = FALSE)
mean(sz)
sigma <- sqrt(sum((sz - mean(sz))^2)/length(sz))
sz_se <- sqrt(sigma^2/30) 
sz_se
quantile(sz)
ci <- quantile(sz, c(0.025, 0.975)) #determines ci without assuming normal distribution
ci 

sy <- sample(d$years_of_education, 30, replace = FALSE)
mean(sy)
sigma <- sqrt(sum((sy - mean(sy))^2)/length(sy))
sy_se <- sqrt(sigma^2/30) 
sy_se
quantile(sy)
ci <- quantile(sy, c(0.025, 0.975)) #determines ci without assuming normal distribution
ci
```
Note: Feels really tedious and I know there's an easier way of doing it but unsure how to shorten the functions


## Question 6
```{r Question 6}
#Now draw 99 more random samples of 30 zombie apocalypse survivors, and calculate the mean for each variable for each of these samples. Together with the first sample you drew, you now have a set of 100 means for each variable (each based on 30 observations), which constitutes a sampling distribution for each variable. What are the means and standard deviations of this distribution of means for each variable? How do the standard deviations of means compare to the standard errors estimated in [5]? What do these sampling distributions look like (a graph might help here)? Are they normally distributed? What about for those variables that you concluded were not originally drawn from a normal distribution?
set.seed(1)
height_99 <- NULL 
for (i in 1:99) {
    height_99[i] <- mean(sample(d$height, size = 30, replace = FALSE))
} #taken from mod 9, gets 99 random samples
all_h <- height_99 + m_sh
mean(all_h)
sd(all_h)

weight_99 <- NULL 
for (i in 1:99) {
    weight_99[i] <- mean(sample(d$weight, size = 30, replace = FALSE))
} #taken from mod 8, gets 99 random samples
all_w <- weight_99 + m_sw
mean(all_w)
sd(all_w)

age_99 <- NULL 
for (i in 1:99) {
    age_99[i] <- mean(sample(d$age, size = 30, replace = FALSE))
} #taken from mod 8, gets 99 random samples
all_a <- age_99 + mean(sa)
mean(all_a)
sd(all_a)

zombkill_99 <- NULL 
for (i in 1:99) {
    zombkill_99[i] <- mean(sample(d$zombies_killed, size = 30, replace = FALSE))
} #taken from mod 8, gets 99 random samples
all_z <- zombkill_99 + mean(sz)
mean(all_z)
sd(all_z)

years_99 <- NULL 
for (i in 1:99) {
    years_99[i] <- mean(sample(d$years_of_education, size = 30, replace = FALSE))
} #taken from mod 8, gets 99 random samples
all_y <- years_99 + mean(sy)
mean(all_y)
sd(all_y)

```

I found that the mean and standard deviation for all the variables had differed alot from the values in question 5. I originally assumed that the mean and sd for the 100 samples should be closer to the normal mean and sd of the whole population but it does not seem to be the case. I believe this issue may be due to my code as something must have gone wrong. The values seem really high compared to the other means and sds. 

