---
title: "weavers_PeerCommentary_Amandafw_03"
author: "Sofia M. Weaver"
date: "`r Sys.Date()`"
output: html_document
---

## Uploading data into R
```{r Uploading Data}
library(curl)
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/zombies.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE) #this is the typical path used in challenges but would give an error for me - "Error in open.connection(file, "rt")" or another error with the final code line
#or 
d <- read.csv("zombies.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE) #this seemed to be more effective 
head(d)
summary(d)
```
# you made a good point about the more effective line of code to read the csv file. it's the same code i used.  - Sofia


## Question 1
```{r Question 1}
#Calculate the population mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, and years of education). NOTE: You will not want to use the built in var() and sd() commands as these are for samples.
names(d)
mean(d$height, na.rm = TRUE) #the mean for heights
h <- d$height
sqrt(sum((h - mean(h))^2)/length(h)) #pop standard deviation is the square root of the pop variance; population variance is the sum of squares divided by # of values

mean(d$weight, na.rm = TRUE)
w <- d$weight
sqrt(sum((w - mean(w))^2)/length(w)) #pop sd

mean(d$age, na.rm = TRUE)
a <- d$age
sqrt(sum((a - mean(a))^2)/length(a)) #pop sd

mean(d$zombies_killed, na.rm = TRUE)
z <- d$zombies_killed
sqrt(sum((z - mean(z))^2)/length(z)) #pop sd

mean(d$years_of_education, na.rm = TRUE)
y <- d$years_of_education
sqrt(sum((y - mean(y))^2)/length(y)) #pop sd
```
# i calculated mean the same way as you, but the way you calculated sd was way more effective than what i did. i eventually figured out the more effective line of code that you used by the end of my Rmd. it was also smart to assign objects. i think it would've made my life easier later in the assignment if i had. - sofia

## Question 2
```{r Question 2}
#Use {ggplot} to make boxplots of each of these variables by gender.
library(ggplot2)
names(d)

h <- ggplot(data = d, aes(x = gender, y = height))
h <- h + geom_boxplot()
h <- h + ylab("Height") + xlab("Gender")

w <- ggplot(data = d, aes(x = gender, y = weight))
w <- w + geom_boxplot()
w <- w + ylab("Weight") + xlab("Gender")

a <- ggplot(data = d, aes(x = gender, y = age))
a <- a + geom_boxplot()
a <- a + ylab("Age") + xlab("Gender")

z <- ggplot(data = d, aes(x = gender, y = zombies_killed))
z <- z + geom_boxplot()
z <- z + ylab("Zombies Killed") + xlab("Gender")

y <- ggplot(data = d, aes(x = gender, y = years_of_education))
y <- y + geom_boxplot()
y <- y + ylab("Years of Education") + xlab("Gender")

library("gridExtra") #package i found to present all the graphs together
grid.arrange(h, w, a, z, y, ncol = 2, nrow = 3) 
```
Source: https://www.geeksforgeeks.org/draw-multiple-ggplot2-plots-side-by-side/
Notes: the graphs look a little condensed though and I'm not too sure how to fix it or whether I should change it so each graph has their own space, however, the knitted code looks really long when I do this
#this looks good. I did it essentially the same way but i have some extra lines of code to add theme and fill the graphs. you can also gropu the graphs together using the gather() function and faceting them. I think it's part of the tidyverse package but don't quote me on that. I didn't use it for this particular asssignment because it condenses *some* of the graphs since it makes them all the same y values. your graphs look really good though. small note - you might want to add some comments to this code. i know we've worked with ggplot() and box plots a good amount but it might help later on if you don't remember - sofia

## Question 3
```{r Question 3}
#Use {ggplot} to make scatterplots of height and weight in relation to age. Do these variables seem to be related? In what way?
p <- ggplot(data = d, aes(x = height, y = age)) #entering the data
p <- p + xlab("Height") + ylab("Age")  #renaming the axis
p <- p + geom_point()  #make a scatterplot
p <- p + geom_smooth(method = "lm", fullrange = TRUE) #adding regression lines

w <- ggplot(data = d, aes(x = weight, y = age)) #entering the data
w <- w + xlab("Weight") + ylab("Age")  #renaming the axis
w <- w + geom_point()  #make a scatterplot
w <- w + geom_smooth(method = "lm", fullrange = TRUE)

grid.arrange(p, w, ncol = 2)
```
There appears to be a more definitive correlation between height and age than weight and age but both variables generally increase with age. Both weight and height appears to be correlated with taller individuals typically having larger weights. 
# these look great. I didn't add a regression line to my graphs but I probably should. - Sofia

## Question 4
```{r Question 4}
#Using histograms and Q-Q plots, check whether the quantitative variables seem to be drawn from a normal distribution. Which seem to be and which do not (hint: not all are drawn from the normal distribution)? For those that are not normal, can you determine from which common distribution they are drawn?
summary(d)
#creating histogram & Q-Q plots with height:
hist(d$height, breaks = seq(from = 50, to = 85, by = 1), probability = TRUE, xlab = "Height", main = "Histogram for Height") #the breaks were determined with summary(d) from the min and the max values
qqnorm(d$height, main = "Q-Q plot for Height")
qqline(d$height, col = "red")

#histogram for weight
hist(d$weight, breaks = seq(from = 85, to = 215, by = 1), probability = TRUE, xlab = "Weight", main = "Histogram for Weight")
qqnorm(d$weight, main = "Q-Q plot for Weight")
qqline(d$weight, col = "red")

#histogram for zombies killed
hist(d$zombies_killed, breaks = seq(from = -5, to = 15, by = 1), probability = TRUE, xlab = "Zombies Killed", main = "Histogram for Zombies Killed")
qqnorm(d$zombies_killed, main = "Q-Q plot for Zombies Killed")
qqline(d$zombies_killed, col = "red")

#histogram for age
hist(d$age, breaks = seq(from = 5, to = 35, by = 1), probability = TRUE, xlab = "Age", main = "Histogram for Age")
qqnorm(d$age, main = "Q-Q plot for Age")
qqline(d$age, col = "red")

#histogram for years of education
hist(d$years_of_education, breaks = seq(from = -5, to = 13, by = 1), probability = TRUE, xlab = "Years of Education", main = "Histogram for Years of Education")
qqnorm(d$years_of_education, main = "Q-Q plot for Years of Education")
qqline(d$years_of_education, col = "red")
```
Height, Weight, and Age appears to be drawn from a normal distribution as they seem to adhere towards the line in the qqplots. In addition, their histograms appear similar to the normal distribution. Years of Education and Zombies Killed did not seem to be drawn from a normal distribution as their qq plots were really different and their histograms appear more similar to poission distributions.

# these look great and I did it essentially the same way, I didn't add breaks however. I'm not sure if they're necessary since all of my graphs look similar to yours? - Sofia

## Question 5
```{r Question 5}
#Now use the sample() function to sample ONE subset of 30 zombie survivors (without replacement) from this population and calculate the mean and sample standard deviation for each variable. Also estimate the standard error for each variable, and construct the 95% confidence interval for each mean. Note that for the variables that are not drawn from the normal distribution, you may need to base your estimate of the CIs on slightly different code than for the normal…
#standard error is the standard deviation of the sampling distribution 
set.seed(1) #makes sure that the sample doesn't change each time
sh <- sample(d$height, 30, replace = FALSE)
m_sh <- mean(sh) #mean
m_sh
sigma <- sqrt(sum((sh - mean(sh))^2)/length(sh)) #sample standard deviation
sigma
sh_se <- sqrt(sigma^2/30) #standard error
sh_se
upper <- mean(sh) + qnorm(0.975, mean = 0, sd = 1) * sh_se
lower <- mean(sh) + qnorm(0.025, mean = 0, sd = 1) * sh_se  #ci with normal distribution
ci <- c(lower, upper)
ci

sw <- sample(d$weight, 30, replace = FALSE)
m_sw <- mean(sw)
m_sw
sigma <- sqrt(sum((sw - mean(sw))^2)/length(sw)) 
sw_se <- sqrt(sigma^2/30) 
sw_se
upper <- mean(sw) + qnorm(0.975, mean = 0, sd = 1) * sw_se
lower <- mean(sw) + qnorm(0.025, mean = 0, sd = 1) * sw_se  #ci with normal distribution
ci <- c(lower, upper)
ci

sa <- sample(d$age, 30, replace = FALSE)
m_sa <- mean(sa)
m_sa
sigma <- sqrt(sum((sa - mean(sa))^2)/length(sa))
sa_se <- sqrt(sigma^2/30) 
sa_se
upper <- mean(sa) + qnorm(0.975, mean = 0, sd = 1) * sa_se
lower <- mean(sa) + qnorm(0.025, mean = 0, sd = 1) * sa_se  #ci with normal distribution
ci <- c(lower, upper)
ci

sz <- sample(d$zombies_killed, 30, replace = FALSE)
m_sz <- mean(sz)
sigma <- sqrt(sum((sz - mean(sz))^2)/length(sz))
sz_se <- sqrt(sigma^2/30) 
sz_se
quantile(sz)
ci <- quantile(sz, c(0.025, 0.975)) #determines ci without assuming normal distribution
ci 

sy <- sample(d$years_of_education, 30, replace = FALSE)
m_sy <- mean(sy)
sigma <- sqrt(sum((sy - mean(sy))^2)/length(sy))
sy_se <- sqrt(sigma^2/30) 
sy_se
quantile(sy)
ci <- quantile(sy, c(0.025, 0.975)) #determines ci without assuming normal distribution
ci
```
Note: Feels really tedious and I know there's an easier way of doing it but unsure how to shorten the functions

#this is exactly how i did it and felt the same way about an easier way. this question was such a pain honestly. great job - sofia 


## Question 6
```{r Question 6}
#Now draw 99 more random samples of 30 zombie apocalypse survivors, and calculate the mean for each variable for each of these samples. Together with the first sample you drew, you now have a set of 100 means for each variable (each based on 30 observations), which constitutes a sampling distribution for each variable. What are the means and standard deviations of this distribution of means for each variable? How do the standard deviations of means compare to the standard errors estimated in [5]? What do these sampling distributions look like (a graph might help here)? Are they normally distributed? What about for those variables that you concluded were not originally drawn from a normal distribution?
set.seed(1)
height_99 <- NULL 
for (i in 1:99) {
    height_99[i] <- mean(sample(d$height, size = 30, replace = FALSE))
} #taken from mod 9, gets 99 random samples
all_h <- c(height_99, m_sh)
mean(all_h)
sd(all_h)

weight_99 <- NULL 
for (i in 1:99) {
    weight_99[i] <- mean(sample(d$weight, size = 30, replace = FALSE))
} #taken from mod 8, gets 99 random samples
all_w <- c(weight_99, m_sw)
mean(all_w)
sd(all_w)

age_99 <- NULL 
for (i in 1:99) {
    age_99[i] <- mean(sample(d$age, size = 30, replace = FALSE))
} #taken from mod 8, gets 99 random samples
all_a <- c(age_99, m_sa)
mean(all_a)
sd(all_a)

zombkill_99 <- NULL 
for (i in 1:99) {
    zombkill_99[i] <- mean(sample(d$zombies_killed, size = 30, replace = FALSE))
} #taken from mod 8, gets 99 random samples
all_z <- c(zombkill_99, m_sz)
mean(all_z)
sd(all_z)

years_99 <- NULL 
for (i in 1:99) {
    years_99[i] <- mean(sample(d$years_of_education, size = 30, replace = FALSE))
} #taken from mod 8, gets 99 random samples
all_y <- c(years_99, m_sy)
mean(all_y)
sd(all_y)
```
I had some issues with this question as originally, I wasn't adding the 99 samples and the previous sample correctly by using the code (Ex: all_w <- weight_99 + m_sw). Instead of giving 100 samples, it kept adding the previous sample onto each value from my loop. However, I received help in class and changed the code to (Ex:all_h <- c(height_99, m_sh)) so it enters everything into one vector so I can calculate the mean. Now, these sd and means are really close to the values calculated from question 5. 

# great job! this question (and the last one) were a challenge for me. we approached our code the same way but I did not have a line of code where i specify i new object with vector c that contained the new sample and your mean value for each particular quantitative value. I tried this out on my code to see if it would make a difference but my numbers were essentially the same. What does this line of code do? - sofia 

# Your code looks great, Amanda! really straightforward for the most part. there were a lot of shortcuts you used that helped you out in the longrun that I think I will implement on my own code. your comments are helpful as well and informative. Great job!